---
title: "Sensitivity/Specificity Analyses - Canine Leishmaniosis"
author: "Marie Ozanne"
date: "December 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
# library(actuar)
library(coda)
library(R2OpenBUGS)
# library(GGally)
library(summarytools)
```

## Exploratory Analyses

```{r exploratory, echo=FALSE}
setwd("G:/My Drive/Research/Sensitivity_Specificity/")
ss_data <- read.csv("DPPReaderData_cleaned_alldata_final.csv", na.strings = c("NA", "", "."))
ss_data <- ss_data[,!(names(ss_data)%in%c("X", "X.1", "X.2", "X.3"))] ## remove columns with comments only
names(ss_data) <- c(names(ss_data)[1:4], "PCR", "PCR_positive",
                    "DPP", "DPP_positive", "ClinicalStatus")

## For analysis purposes - treat all observations as independent and 'Timepoint' is irrelevant
## Clinical Status (based on 2 or more clinical symptoms): N=negative, A=asymptomatic, S=symptomatic

ss_data[which(ss_data$ClinicalStatus=="AS"),]$ClinicalStatus <- "A" ## replace AS with A for consistency
ss_data[which(ss_data$ClinicalStatus=="SY"),]$ClinicalStatus <- "S" ## replace SY with S for consistency

## drop unused levels from Clinical Status
ss_data$ClinicalStatus <- droplevels(ss_data$ClinicalStatus)

## "F" and "M" erroneously places in Age column
ss_data$Age[which(ss_data$Age=="F")] <- NA
ss_data$Age[which(ss_data$Age=="M")] <- NA
ss_data$Age <- as.numeric(ss_data$Age)

## One pup has PCR as positive - moving this indication to the PCR_positive
## column, which currently has an NA
ss_data[is.na(ss_data$PCR_positive),]$PCR_positive <- "Positive"

## Standardize the levels for PCR_positive
## Current levels: > levels(ss_data$PCR_positive)
#  [1] "BL"                         "BL (single 1:10 well only)" "BL below threshold"   
#  [4] "N"                          "Neg"                        "negative"     
#  [7] "Negative"                   "P"                          "pos"       
# [10] "Pos"                        "positive"                   "Positive"  
### Assume all BL (below limits) are "Negative"
ss_data[ss_data$PCR_positive=="BL",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="BL (single 1:10 well only)",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="BL below threshold",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="N",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="Neg",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="negative",]$PCR_positive <- "Negative"

### Standardize all positive entries to "Positive"
ss_data[ss_data$PCR_positive=="P",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="pos",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="Pos",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="positive",]$PCR_positive <- "Positive"

## drop unused levels from PCR_positive
ss_data$PCR_positive <- droplevels(ss_data$PCR_positive)

## One pup had "positive" in the PCR column, which should be numeric
## replace with NA - no numerical value available
ss_data$PCR[which(ss_data$PCR=="positive")] <- NA

## Change PCR to numeric
ss_data$PCR <- as.numeric(ss_data$PCR)

## Print summary of data
summary(ss_data[,!(names(ss_data) %in% c("Timepoint", "ID"))])

## Create new DPP_positive variable based on reader cutoffs (>9.9 is positive)
ss_data$DPP_positive2 <- ifelse(ss_data$DPP > 9.9,
                                c("Positive"), c("Negative"))
```

## Model 1:

This model is designed to let us differentiate between negative and positive (asymptomatic or symptomatic as determined by physical exam information - 2 or more symptoms) dogs based on positive/negative PCR and DPP and associated sensitivity and specificity. This assumes the two tests are positively correlated.

### Model Setup

#### Notation:

* $Y^{PCR}_i$: PCR test result for observation $i$
* $Y^{DPP}_i$: DPP test result for observation $i$ (DPP reader \underline{or} DPP positive/negative)
* $\mathbf{Y}_i=(Y^{DPP}_i, Y^{PCR}_i)$
* $D_i$: disease status for observation $i$ (positive/negative)
* $S^{DPP}=P(DPP^+|D)$: sensitivity of DPP test
* $S^{PCR}=P(PCR^+|D)$: sensitivity of PCR test
* $C^{DPP}=P(DPP^-|D^c):$ specificity of DPP test
* $C^{PCR}=P(PCR^-|D^c):$ specificity of PCR test


#### Data Model:

$$
\begin{aligned}
Y^{DPP}_i|\pi^{DPP}_i, D_i&\sim Bernoulli(\pi^{DPP}_i)
\end{aligned}
$$


$$
\begin{aligned}
Y^{PCR}_i|\pi^{PCR}_i, D_i&\sim Bernoulli(\pi^{PCR}_i)
\end{aligned}
$$


#### Process Model:

DPP
$$
\begin{aligned}
\pi_i^{DPP}&=P(DPP^+\cap PCR^+ \cap D_i)+P(DPP^+\cap PCR^+\cap D_i^c)\\
& \ \ \ + P(DPP^+\cap PCR^-\cap D_i) + P(DPP^+\cap PCR^-\cap D_i^c)\\
&=P(DPP^+\cap PCR^+ | D_i)P(D_i)+P(DPP^+\cap PCR^+|D_i^c)P(D_i^c)\\
& \ \ \ + P(DPP^+\cap PCR^-|D_i)P(D_i) + P(DPP^+\cap PCR^-|D_i^c)P(D_i^c)\\
&=((S^{DPP}\times S^{PCR}+covs_{12})+(S^{DPP}\times(1-S^{PCR})-covs_{12}))\times P(D_i)\\
& \ \ \ + [((1-C^{DPP})\times (1-C^{PCR})+covc_{12})+((1-C^{DPP})\times(C^{PCR})-covc_{12})]\times P(D_i^c)\\
\end{aligned}
$$

PCR
$$
\begin{aligned}
\pi_i^{PCR}&=P(DPP^+\cap PCR^+\cap D_i)+P(DPP^+\cap PCR^+\cap D_i^c)\\
& \ \ \ +P(DPP^-\cap PCR^+\cap D_i)+P(DPP^-\cap PCR^+\cap D_i^c)\\
&=P(DPP^+\cap PCR^+|D_i)P(D_i)+P(DPP^+\cap PCR^+| D_i^c)P(D_i^c)\\
& \ \ \ +P(DPP^-\cap PCR^+|D_i)P(D_i)+P(DPP^-\cap PCR^+|D_i^c)P(D_i^c)
\end{aligned}
$$

Disease
$$
D_i|\boldsymbol{\beta}\sim Bernoulli(\pi^D_i)
$$
where 

$$\pi^D_i=P(D_i)=\frac{\exp\left\{\epsilon_i + X_i\boldsymbol{\beta} \right\}}{1+\exp\left\{\epsilon_i+X_i\boldsymbol{\beta} \right\}}; \ X_i=[1, \mathbf{1}_{Sex_i=M}, Age_i]$$.


#### Parameter Model:
$$
\boldsymbol{\beta}\sim MVN(\boldsymbol{\mu}_\beta, \Sigma_\beta).
$$

### Model Fit:

The model was fit using OpenBUGS via the R2OpenBUGS package in R. 

```{r OpenBUGS Model 1, echo=FALSE, results="hide"}
## ranges of sensitivities and specificities
sens.pcr.range <- c(0.839, 0.990)
sens.dpp.range <- c(0.832, 0.930)
spec.pcr.range <- c(0.871, 0.970)
spec.dpp.range <- c(0.682, 0.951)

sens.pcr <- mean(sens.pcr.range)
sens.dpp <- mean(sens.dpp.range)
spec.pcr <- mean(spec.pcr.range)
spec.dpp <- mean(spec.dpp.range)

prev.range <- c(0.05,0.10)
prev <- mean(prev.range)

## Omit missing values
ss_data_nao <- na.omit(ss_data[,names(ss_data) %in% c("Sex", "Age", "DPP",
                                                      "ClinicalStatus",
                                                      "DPP_positive",
                                                      "PCR_positive",
                                                      "DPP_positive2")])
set.seed(123124)
train_idx <- sample(1:nrow(ss_data_nao), 0.8*nrow(ss_data_nao))
# ss_data2 <- ss_data_nao[train_idx,]
ss_data2 <- ss_data_nao[ss_data_nao$ClinicalStatus!="A",]

nind <- nrow(ss_data2)

y.dpp1.pcr1 <- 1*(ss_data2$DPP_positive2=="Positive" & ss_data2$PCR_positive=="Positive")
y.dpp1.pcr0 <- 1*(ss_data2$DPP_positive2=="Positive" & ss_data2$PCR_positive=="Negative")
y.dpp0.pcr1 <- 1*(ss_data2$DPP_positive2=="Negative" & ss_data2$PCR_positive=="Positive")
y.dpp0.pcr0 <- 1*(ss_data2$DPP_positive2=="Negatove" & ss_data2$PCR_positive=="Negative")

Sex <- 1*(ss_data2$Sex=="M")
Age <- ss_data2$Age
Age2 <- as.numeric(Age)
DPP <- ss_data2$DPP

a_prev <- prev/(1-prev)
b_prev <- 1

a_s_dpp <- sens.dpp/(1-sens.dpp)
b_s_dpp <- 1
a_c_dpp <- spec.dpp/(1-spec.dpp)
b_c_dpp <- 1

a_lambda.D <- 1
b_lambda.D <- 1
a_gamma.D <- 1
b_gamma.D <- 1
a_lambda.Dbar <- 1
b_lambda.Dbar <- 1
a_gamma.Dbar <- 1
b_gamma.Dbar <- 1

data <- list(nind=nind,
             y.dpp1.pcr1=y.dpp1.pcr1,
             y.dpp1.pcr0=y.dpp1.pcr0,
             y.dpp0.pcr1=y.dpp0.pcr1,
             y.dpp0.pcr0=y.dpp0.pcr0,
             Sex=Sex,
             Age=Age2,
             a_prev=a_prev,
             b_prev=b_prev,
             a_s_dpp=a_s_dpp,
             b_s_dpp=b_s_dpp,
             a_c_dpp=a_c_dpp,
             b_c_dpp=b_c_dpp,
             a_lambda.D=a_lambda.D,
             b_lambda.D=b_lambda.D,
             a_gamma.D=a_gamma.D,
             b_gamma.D=b_gamma.D,
             a_lambda.Dbar=a_lambda.Dbar,
             b_lambda.Dbar=b_lambda.Dbar,
             a_gamma.Dbar=a_gamma.Dbar,
             b_gamma.Dbar=b_gamma.Dbar)

## Initialize all prior values
inits <- function(){
  list(pi=rbeta(1,a_prev,b_prev),
       eta.dpp=rbeta(1,a_s_dpp,b_s_dpp),
       theta.dpp=rbeta(1,a_c_dpp,b_c_dpp),
       lambda.D=rbeta(1,a_lambda.D, b_lambda.D),
       gamma.D=rbeta(1,a_gamma.D, b_gamma.D),
       lambda.Dbar=rbeta(1,a_lambda.Dbar, b_lambda.Dbar),
       gamma.Dbar=rbeta(1,a_gamma.Dbar, b_lambda.Dbar),
       b1=rnorm(1,0,2),
       b2=rnorm(1,0,2),
       b3=rnorm(1,0,2),
       eps=rnorm(nind,0,2))
}

## Start MCMC simulation
sens_spec.sim <- bugs(data=data, 
                      inits=inits, ## FIX THIS LINE
                      model.file="OpenBUGS_S_Sp_NSonly_Georgiadis_2003.txt",
                      parameters=c("b1","b2","b3","eps","pi.D",
                                   "rho.D", "rho.Dbar"),
                      n.chains=2, n.iter=5000, n.burnin=1000, 
                      codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject1 <- read.bugs(sens_spec.sim)
model1_df <- do.call(rbind.data.frame, codaobject1)

## Graphial summaries of posterior distribtuions
par(mfrow=c(1,3))
hist(model1_df$b1, main="Intercept (b0)", xlab="posterior draws")
abline(v=mean(model1_df$b1), lty="dashed", col="red")
hist(model1_df$b2, main="Sex=Male (b1)", xlab="posterior draws")
abline(v=mean(model1_df$b2), lty="dashed", col="red")
hist(model1_df$b3, main="Age (b2)", xlab="posterior draws")
abline(v=mean(model1_df$b3), lty="dashed", col="red")
dev.off()

par(mfrow=c(1,2))
hist(model1_df$rho.D, main="Correlation D", xlab="posterior draws")
abline(v=mean(model1_df$rho.D), lty="dashed", col="red")
hist(model1_df$rho.Dbar, main="Correlation Dbar", xlab="posterior draws")
abline(v=mean(model1_df$rho.Dbar), lty="dashed", col="red")
dev.off()

## Numeric summaries of posterior distributions
# boxplot(model1_df[,!(names(model1_df) %in% c("deviance"))])

```

### Model Predictions

```{r Model Predictions Model 1, echo=FALSE}
## Want to predict disease state using our model
## Posterior draws
set.seed(123123)
nn <- 500

Beta_m1 <- model1_df[,names(model1_df) %in% c("b1","b2","b3")]
eps_m1 <- as.data.frame(model1_df[,grepl(names(model1_df), pattern="eps", fixed=TRUE)])
pi.D_m1 <- as.data.frame(model1_df[,grepl(names(model1_df), pattern="pi.D", fixed=TRUE)])
pi.D_m1 <- pi.D_m1[5001:nrow(pi.D_m1),]

## define test set
test_set1 <- ss_data_nao[ss_data_nao$ClinicalStatus!="A",]

model1_PD <- matrix(data=NA, nrow=nrow(test_set1),ncol=nn)

## Summarize compartment membership probabilities with median and CIs
tmp_PD <- apply(pi.D_m1, 2, quantile, c(0.025,0.50,0.975))

pred_m1_df <- data.frame(Age=test_set1$Age,
                         Sex=test_set1$Sex,
                         Prob_D_median=t(tmp_PD)[,2],
                         Prob_D_lb=t(tmp_PD)[,1],
                         Prob_D_ub=t(tmp_PD)[,3],
                         ClinicalStatus=test_set1$ClinicalStatus,
                         PCR_positive=test_set1$PCR_positive,
                         DPP_positive=test_set1$DPP_positive)  

print("Clinical Status == N")
summary(pred_m1_df[pred_m1_df$ClinicalStatus=="N",])

print("Clinical Status == S")
summary(pred_m1_df[pred_m1_df$ClinicalStatus=="S",])

## check correspondance with tests
print("PCR Positive and DPP Positive")
summary(pred_m1_df[pred_m1_df$PCR_positive=="Positive" & pred_m1_df$DPP_positive=="Positive",])

print("PCR Positive and DPP Negative -- NOBODY HERE")
# summary(pred_m1_df[pred_m1_df$PCR_positive=="Positive" & pred_m1_df$DPP_positive=="Negative",])

print("PCR Negative and DPP Positive")
summary(pred_m1_df[pred_m1_df$PCR_positive=="Negative" & pred_m1_df$DPP_positive=="Positive",])

print("PCR Negative and DPP Negative")
summary(pred_m1_df[pred_m1_df$PCR_positive=="Negative" & pred_m1_df$DPP_positive=="Negative",])

```
