---
title: "Sensitivity/Specificity Analyses - Canine Leishmaniosis"
author: "Marie Ozanne"
date: "March 3, 2020"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
# library(actuar)
library(coda)
library(egg)
library(ggplot2)
library(R2OpenBUGS)
# library(GGally)
library(summarytools)
```

# Exploratory Analyses

```{r exploratory, echo=FALSE}
setwd("G:/My Drive/Research/Sensitivity_Specificity/")
ss_data <- read.csv("Data/DPPReaderData_cleaned_alldata_final.csv", na.strings = c("NA", "", "."))
ss_data <- ss_data[,!(names(ss_data)%in%c("X", "X.1", "X.2", "X.3"))] ## remove columns with comments only
names(ss_data) <- c(names(ss_data)[1:4], "PCR", "PCR_positive",
                    "DPP", "DPP_positive", "ClinicalStatus")

## For analysis purposes - treat all observations as independent and 'Timepoint' is irrelevant
## Clinical Status (based on 2 or more clinical symptoms): N=negative, A=asymptomatic, S=symptomatic

ss_data[which(ss_data$ClinicalStatus=="AS"),]$ClinicalStatus <- "A" ## replace AS with A for consistency
ss_data[which(ss_data$ClinicalStatus=="SY"),]$ClinicalStatus <- "S" ## replace SY with S for consistency

## drop unused levels from Clinical Status
ss_data$ClinicalStatus <- droplevels(ss_data$ClinicalStatus)

## "F" and "M" erroneously places in Age column
ss_data$Age[which(ss_data$Age=="F")] <- NA
ss_data$Age[which(ss_data$Age=="M")] <- NA
ss_data$Age <- as.numeric(ss_data$Age)

## One pup has PCR as positive - moving this indication to the PCR_positive
## column, which currently has an NA
ss_data[is.na(ss_data$PCR_positive),]$PCR_positive <- "Positive"

## Standardize the levels for PCR_positive
## Current levels: > levels(ss_data$PCR_positive)
#  [1] "BL"                         "BL (single 1:10 well only)" "BL below threshold"   
#  [4] "N"                          "Neg"                        "negative"     
#  [7] "Negative"                   "P"                          "pos"       
# [10] "Pos"                        "positive"                   "Positive"  
### Assume all BL (below limits) are "Negative"
ss_data[ss_data$PCR_positive=="BL",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="BL (single 1:10 well only)",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="BL below threshold",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="N",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="Neg",]$PCR_positive <- "Negative"
ss_data[ss_data$PCR_positive=="negative",]$PCR_positive <- "Negative"

### Standardize all positive entries to "Positive"
ss_data[ss_data$PCR_positive=="P",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="pos",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="Pos",]$PCR_positive <- "Positive"
ss_data[ss_data$PCR_positive=="positive",]$PCR_positive <- "Positive"

## drop unused levels from PCR_positive
ss_data$PCR_positive <- droplevels(ss_data$PCR_positive)

## One pup had "positive" in the PCR column, which should be numeric
## replace with NA - no numerical value available
ss_data$PCR[which(ss_data$PCR=="positive")] <- NA

## Change PCR to numeric
ss_data$PCR <- as.numeric(ss_data$PCR)

## Print summary of data
# summary(ss_data[,!(names(ss_data) %in% c("Timepoint", "ID"))])

## Create new DPP_positive variable based on reader cutoffs (>9.9 is positive)
ss_data$DPP_positive2 <- ifelse(ss_data$DPP > 9.9,
                                c("Positive"), c("Negative"))

## Omit missing values
ss_data_nao <- na.omit(ss_data[,names(ss_data) %in% c("Sex", "Age", "DPP",
                                                      "ClinicalStatus",
                                                      "DPP_positive",
                                                      "PCR_positive",
                                                      "DPP_positive2")])

## Create contingency table for data we are using:
ss_data2 <- ss_data_nao[ss_data_nao$ClinicalStatus!="A",]
ss_data2$DPP_positive2 <- as.factor(ss_data2$DPP_positive2)
table(PCR=ss_data2$PCR_positive, DPP=ss_data2$DPP_positive2)

## Include variable for diagnostically positive (positive on at least one test)
ss_data2$Diagnostically_positive <- rep(NA, nrow(ss_data2))
ss_data2[ss_data2$PCR_positive == "Positive" | ss_data2$DPP_positive2 == "Positive",]$Diagnostically_positive <- "Positive"
ss_data2[ss_data2$PCR_positive == "Negative" & ss_data2$DPP_positive2 == "Negative",]$Diagnostically_positive <- "Negative"
```

# Models

Angela's paper (Toepp et al., 2019, https://doi.org/10.1371/journal.pntd.0007058) uses logistic regression, with age, sex, and variables that have to do with diagnostic tests as explanatory variables. They are something like this:  

**Model A 1:** $logit(\pi_k)=\beta_0+\beta_1Age_k + \beta_2Sex_k + \beta_3 Y_k$, where $Y_k$ is diagnostically positive (as defined in Model 1 below), but for the mom and $\pi_k$ is the probability of disease for individual $k$

**Model A 2:** $logit(\pi_k)=\beta_0+\beta_1Age_k + \beta_2Sex_k + \beta_3 T_{1k} + \beta_4T_{2k}$, where $T_{jk}$ is the result for Test $j$ (as defined in Model 1 below), but for the mom and $\pi_k$ is the probability of disease for individual $k$

Note, these models were fit with a log link function, presumably so that relative risks could be recovered?

We plan to evaluate similar models for our data and to then incorporate sensitivity and specificity of the tests into these models. Then we will compare the model performance to that of other methods. Hopefully we will see an improvement/some details that we miss when we do not include the sensitivity and specificity for the tests.

In all these models, we will assume that the observations are independent.

## Separate Diagnostic Tests in Data Model:

### Data Model

$$
T_{1k}|D_k \sim Bernoulli(\pi_{1k})
$$
where $\pi_{1k}=P(T_{1k}=1|D_k)=D_k\times \underbrace{P(T_{1k}=1|D_k=1)}_{sensitivity}+(1-D_k)\times \underbrace{(1-P(T_{1k}=0|D_k=0))}_{1-specificity}$.
$$
T_{2k}|D_k \sim Bernoulli(\pi_{2k})
$$
where $\pi_{2k}=P(T_{2k}=1|D_k)=D_k\times \underbrace{P(T_{2k}=1|D_k=1)}_{sensitivity}+(1-D_k)\times \underbrace{(1-P(T_{2k}=0|D_k=0))}_{1-specificity}$.

### Process Model

$$
D_k \sim Bernoulli(\delta_k)
$$
where $\delta_k=logit^{-1}(logit(\rho)+\mathbf{x}_k^T\boldsymbol{\beta}+\epsilon_k)$, and $\epsilon_k$ is a random effect for individual $k$.

### Prior Model

#### Prevalence
$$
logit(\rho) \sim Normal(\mu_{\rho}, \tau^2)
$$
where $\mu_\rho=logit^{-1}(\rho^*)$ and $\tau^2=100$.

<!-- We have a range for the prevalence of (0.05, 0.10). This corresponds to a range of (`r round(log(0.05)/(1-log(0.05)),4)`, `r round(log(0.10)/(1-log(0.10)),4)`) on the logit scale.  -->

<!-- ```{r} -->
<!-- hist(rnorm(10000, mean = log(0.075)/(1-log(0.075)), sd = 0.03), -->
<!--      main="Logit prevalence prior distribution histogram", -->
<!--      xlab="logit(prevalence)") -->
<!-- ``` -->

<!-- Note, the prevalence in this population may be higher - these are exposed dogs in the United States in our hunting hound population. We may want to change this, but we also want to generalize this to the larger canine population - to Brazil, if possible. -->

#### Random intercept

$$
\boldsymbol{\epsilon}\sim Normal(0,5000)
$$

## OpenBUGS Implementation (a): 

  - no individual fixed effects
  - sens=spec=0.999
  
```{r Open Bugs setup_1a}
## ranges of sensitivities and specificities
sens.pcr <- 0.999
sens.dpp <- 0.999
spec.pcr <- 0.999
spec.dpp <- 0.999

## range of prevalence for visceral leishmaniasis
prev.range <- c(0.05,0.10)
prev <- mean(prev.range)
```

```{r Open Bugs Model 1a, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp <- 1*(ss_data2$DPP_positive2=="Positive")
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
Ind_Age <- 1*(Age >= 6)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp=y.dpp,
               sens.pcr=sens.pcr,
               spec.pcr=spec.pcr,
               sens.dpp=sens.dpp,
               spec.dpp=spec.dpp
             )
  
## Initialize all prior values
inits <- function(){
    list(b1=rnorm(0,1),
         eps=rnorm(nind,0,5000))
}
  
  ## Start MCMC simulation
sens_spec.sim_m1a <- bugs(data=data, 
                          inits=inits, ## FIX THIS LINE
                          model.file="Model1a_separate_tests_GB.txt",
                          parameters=c("b1",
                                       "D", 
                                       "eps"),
                          n.chains=3, n.iter=30000, n.burnin=10000, 
                          codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_m1a <- read.bugs(sens_spec.sim_m1a)
model1a_df <- do.call(rbind.data.frame, codaobject_m1a)
```

- Mixing seems to be much improved (for those individuals where the trace plots is anything other than 0) over previous model.

### Posterior Distributions (a)

```{r}
# par(mfrow=c(2,2))
## Graphial summaries of posterior distributions
# hist(exp(model1a_df$lpi)/(1+exp(model1a_df$lpi)), main="prevalence", xlab="posterior draws")
# abline(v=mean(exp(model1a_df$lpi)/(1+exp(model1a_df$lpi))), lty="dashed", col="red")

hist(exp(model1a_df$b1)/(1+exp(model1a_df$b1)), main="Intercept", xlab="posterior draws")
abline(v=mean(model1a_df$b1), lty="dashed", col="red")
# hist(model1a_df$b2, main="Sex", xlab="posterior draws")
# abline(v=mean(model1a_df$b2), lty="dashed", col="red")
# hist(model1a_df$b3, main="Age/Sex Interaction", xlab="posterior draws")
# abline(v=mean(model1a_df$b3), lty="dashed", col="red")

summary(exp(model1a_df$b1)/(1+exp(model1a_df$b1)))
```

### Disease State Classification (a)

```{r}
## Set up storage for model results
pred_df_m1a <- data.frame(obs=1:nind,
                         D=rep(NA,nind), ## average estimate
                         SD=rep(NA,nind),
                         LB=rep(NA,nind), ## 2.5th percentile
                         UB=rep(NA,nind), ## 97.5th percentile
                         model_assignment=rep(NA,nind),
                         Clinical_status=ss_data2$ClinicalStatus,
                         Diagnostic_status=ss_data2$Diagnostically_positive)

## Calculate probabilities of compartment membership for each posterior draw
pred_df_m1a$D <- apply(model1a_df[,grep("D", names(model1a_df))], 2, mean)
pred_df_m1a$SD <- apply(model1a_df[,grep("D", names(model1a_df))], 2, sd)
pred_df_m1a$LB <- apply(model1a_df[,grep("D", names(model1a_df))], 2, 
                       quantile, probs=0.025)
pred_df_m1a$UB <- apply(model1a_df[,grep("D", names(model1a_df))], 2, 
                       quantile, probs=0.975)


summary(pred_df_m1a)

## Apply a cut off of point estimate of 0.5; if pi.D > 0.5, classify as S (symptomatic), otherwise as N;
## Summarize in a table (clinical status versus diagnostic status)
table(pred_df_m1a[pred_df_m1a$D > 0.5,]$Clinical_status, 
      pred_df_m1a[pred_df_m1a$D > 0.5,]$Diagnostic_status)

## Print summary table of clinical status versus diagnostic status from the original data
table(pred_df_m1a$Clinical_status, pred_df_m1a$Diagnostic_status)

## boxplots
p1 <- (ggplot(data=pred_df_m1a, aes(x=Diagnostic_status, y=D)) 
       + geom_boxplot()
       + theme_bw())
p2 <- (ggplot(data=pred_df_m1a, aes(x=Clinical_status, y=D)) 
       + geom_boxplot()
       + theme_bw())

ggarrange(p1,p2, nrow=1)
```


## OpenBUGS Implementation (b)

Same as (a), but without fixing the sensitivities and specifities to be 0.999.

```{r Open Bugs setup_1b}
## ranges of sensitivities and specificities
sens.pcr.range <- c(0.839, 0.990)
sens.dpp.range <- c(0.832, 0.930)
spec.pcr.range <- c(0.871, 0.970)
spec.dpp.range <- c(0.682, 0.951)

sens.pcr <- mean(sens.pcr.range)
sens.dpp <- mean(sens.dpp.range)
spec.pcr <- mean(spec.pcr.range)
spec.dpp <- mean(spec.dpp.range)

## range of prevalence for visceral leishmaniasis
prev.range <- c(0.05,0.10)
prev <- mean(prev.range)
```

```{r Open Bugs Model 1b, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp <- 1*(ss_data2$DPP_positive2=="Positive")
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
Ind_Age <- 1*(Age >= 6)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp=y.dpp,
               sens.pcr=sens.pcr,
               spec.pcr=spec.pcr,
               sens.dpp=sens.dpp,
               spec.dpp=spec.dpp
             )
  
## Initialize all prior values
inits <- function(){
    list(b1=rnorm(0,1),
         eps=rnorm(nind,0,5000))
}

## Start MCMC simulation
sens_spec.sim_m1b <- bugs(data=data, 
                       inits=inits, ## FIX THIS LINE
                       model.file="Model1a_separate_tests_GB.txt",
                       parameters=c("b1", 
                                    "D", 
                                    "eps"),
                       n.chains=3, n.iter=30000, n.burnin=10000, 
                       codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_m1b <- read.bugs(sens_spec.sim_m1b)
model1b_df <- do.call(rbind.data.frame, codaobject_m1b)
```

### Posterior Distributions (b)

```{r}
# par(mfrow=c(2,2))
## Graphial summaries of posterior distributions
hist(exp(model1b_df$b1)/(1+exp(model1b_df$b1)), main="prevalence", xlab="posterior draws")
abline(v=mean(exp(model1b_df$b1)/(1+exp(model1b_df$b1))), lty="dashed", col="red")

# hist(model1b_df$b1, main="Age", xlab="posterior draws")
# abline(v=mean(model1b_df$b1), lty="dashed", col="red")
# hist(model1b_df$b2, main="Sex", xlab="posterior draws")
# abline(v=mean(model1b_df$b2), lty="dashed", col="red")
# hist(model1b_df$b3, main="Age/Sex Interaction", xlab="posterior draws")
# abline(v=mean(model1b_df$b3), lty="dashed", col="red")
```

### Disease State Classification (b)

```{r}
## Set up storage for model results
pred_df_m1b <- data.frame(obs=1:nind,
                         D=rep(NA,nind), ## average estimate
                         SD=rep(NA,nind),
                         LB=rep(NA,nind), ## 2.5th percentile
                         UB=rep(NA,nind), ## 97.5th percentile
                         model_assignment=rep(NA,nind),
                         Clinical_status=ss_data2$ClinicalStatus,
                         Diagnostic_status=ss_data2$Diagnostically_positive)

## Calculate probabilities of compartment membership for each posterior draw
pred_df_m1b$D <- apply(model1b_df[,grep("D", names(model1b_df))], 2, mean)
pred_df_m1b$SD <- apply(model1b_df[,grep("D", names(model1b_df))], 2, sd)
pred_df_m1b$LB <- apply(model1b_df[,grep("D", names(model1b_df))], 2, 
                       quantile, probs=0.025)
pred_df_m1b$UB <- apply(model1b_df[,grep("D", names(model1b_df))], 2, 
                       quantile, probs=0.975)


summary(pred_df_m1b)

## Apply a cut off of point estimate of 0.5; if pi.D > 0.5, classify as S (symptomatic), otherwise as N;
## Summarize in a table (clinical status versus diagnostic status)
table(pred_df_m1b[pred_df_m1b$D > 0.5,]$Clinical_status, 
      pred_df_m1b[pred_df_m1b$D > 0.5,]$Diagnostic_status)

## Print summary table of clinical status versus diagnostic status from the original data
table(pred_df_m1b$Clinical_status, pred_df_m1b$Diagnostic_status)

## boxplots
p1 <- (ggplot(data=pred_df_m1b, aes(x=Diagnostic_status, y=D)) 
       + geom_boxplot()
       + theme_bw())
p2 <- (ggplot(data=pred_df_m1b, aes(x=Clinical_status, y=D)) 
       + geom_boxplot()
       + theme_bw())

ggarrange(p1,p2, nrow=1)
```

- On the plus side, the mixing looks much better with this model.

## OpenBUGS implementation (c)

#### Linear predictors

The regression parameters are $\boldsymbol{\beta}=(\beta_{age},\beta_{sex},\beta_{age*sex})^T$;

$$
  \boldsymbol{\beta}\sim Normal(\boldsymbol{\mu}_\beta,\boldsymbol{\Sigma}_\beta)
$$
  where $\boldsymbol{\mu}_\beta=\mathbf{0}$ and $\boldsymbol{\Sigma}_\beta=\mathbf{I}$ in our code. 
  
```{r}
## ranges of sensitivities and specificities
sens.pcr.range <- 0.999
sens.dpp.range <- 0.999
spec.pcr.range <- 0.999
spec.dpp.range <- 0.999

  
## range of prevalence for visceral leishmaniasis
prev.range <- c(0.05,0.10)
prev <- mean(prev.range)
```

### Posterior Distributions (c)

```{r Open Bugs Model 1c, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp <- 1*(ss_data2$DPP_positive2=="Positive")
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
Ind_Age <- 1*(Age >= 6)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp=y.dpp,
               Sex=Sex2,
               Age=Age,
               Ind_Age=Ind_Age,
               sens.pcr=sens.pcr,
               spec.pcr=spec.pcr,
               sens.dpp=sens.dpp,
               spec.dpp=spec.dpp
             )
  
## Initialize all prior values
inits <- function(){
    list(b1=rnorm(0,1),
         b2=rnorm(0,1),
         eps=rnorm(nind,0,5000))
}

## Start MCMC simulation
sens_spec.sim_m1c <- bugs(data=data, 
                       inits=inits, ## FIX THIS LINE
                       model.file="Model1b_separate_tests_GB.txt", ## update model
                       parameters=c("b1", "b2", "b3", "b4",
                                    "D", 
                                    "eps"),
                       n.chains=3, n.iter=30000, n.burnin=10000, 
                       codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_m1c <- read.bugs(sens_spec.sim_m1c)
model1c_df <- do.call(rbind.data.frame, codaobject_m1c)
```

```{r}
par(mfrow=c(2,2))
## Graphial summaries of posterior distributions
hist(exp(model1c_df$b1)/(1+exp(model1c_df$b1)), main="prevalence", xlab="posterior draws")
abline(v=mean(exp(model1c_df$b1)/(1+exp(model1c_df$b1))), lty="dashed", col="red")

hist(model1c_df$b2, main="Age", xlab="posterior draws")
abline(v=mean(model1c_df$b2), lty="dashed", col="red")
hist(model1c_df$b3, main="Sex", xlab="posterior draws")
abline(v=mean(model1c_df$b3), lty="dashed", col="red")
hist(model1c_df$b4, main="Age/Sex Interaction", xlab="posterior draws")
abline(v=mean(model1c_df$b4), lty="dashed", col="red")
```

### Disease State Classification (c)

```{r}
## Set up storage for model results
pred_df_m1c <- data.frame(obs=1:nind,
                         D=rep(NA,nind), ## average estimate
                         SD=rep(NA,nind),
                         LB=rep(NA,nind), ## 2.5th percentile
                         UB=rep(NA,nind), ## 97.5th percentile
                         model_assignment=rep(NA,nind),
                         Clinical_status=ss_data2$ClinicalStatus,
                         Diagnostic_status=ss_data2$Diagnostically_positive)

## Calculate probabilities of compartment membership for each posterior draw
pred_df_m1c$D <- apply(model1c_df[,grep("D", names(model1c_df))], 2, mean)
pred_df_m1c$SD <- apply(model1c_df[,grep("D", names(model1c_df))], 2, sd)
pred_df_m1c$LB <- apply(model1c_df[,grep("D", names(model1c_df))], 2, 
                       quantile, probs=0.025)
pred_df_m1c$UB <- apply(model1c_df[,grep("D", names(model1c_df))], 2, 
                       quantile, probs=0.975)


summary(pred_df_m1c)

## Apply a cut off of point estimate of 0.5; if pi.D > 0.5, classify as S (symptomatic), otherwise as N;
## Summarize in a table (clinical status versus diagnostic status)
table(pred_df_m1c[pred_df_m1c$D > 0.5,]$Clinical_status, 
      pred_df_m1c[pred_df_m1c$D > 0.5,]$Diagnostic_status)

## Print summary table of clinical status versus diagnostic status from the original data
table(pred_df_m1c$Clinical_status, pred_df_m1c$Diagnostic_status)

## boxplots
p1 <- (ggplot(data=pred_df_m1c, aes(x=Diagnostic_status, y=D)) 
       + geom_boxplot()
       + theme_bw())
p2 <- (ggplot(data=pred_df_m1c, aes(x=Clinical_status, y=D)) 
       + geom_boxplot()
       + theme_bw())

ggarrange(p1,p2, nrow=1)
```


## OpenBUGS implementation (d)













#### Linear predictors

The regression parameters are $\boldsymbol{\beta}=(\beta_{age},\beta_{sex},\beta_{age*sex})^T$;

$$
  \boldsymbol{\beta}\sim Normal(\boldsymbol{\mu}_\beta,\boldsymbol{\Sigma}_\beta)
$$
  where $\boldsymbol{\mu}_\beta=\mathbf{0}$ and $\boldsymbol{\Sigma}_\beta=\mathbf{I}$ in our code. 
  
  
```{r Open Bugs setup2 other}
  ## ranges of sensitivities and specificities
  sens.pcr.range <- c(0.839, 0.990)
  sens.dpp.range <- c(0.832, 0.930)
  spec.pcr.range <- c(0.871, 0.970)
  spec.dpp.range <- c(0.682, 0.951)
  
  sens.pcr <- mean(sens.pcr.range)
  sens.dpp <- mean(sens.dpp.range)
  spec.pcr <- mean(spec.pcr.range)
  spec.dpp <- mean(spec.dpp.range)
  
  ## range of prevalence for visceral leishmaniasis
  prev.range <- c(0.05,0.10)
  prev <- mean(prev.range)
```
  
## OpenBUGS Model 1 Implementation
  
```{r Open Bugs Model 1 other, include=FALSE, message=FALSE}
  ## Specify data
  y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
  y.dpp <- 1*(ss_data2$DPP_positive2=="Positive")
  
  nind <- nrow(ss_data2)
  
  Sex2 <- 1*(ss_data2$Sex=="M")
  Age <- as.numeric(ss_data2$Age)
  
  data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp=y.dpp,
               Sex=Sex2,
               Age=Age,
               sens.pcr=sens.pcr,
               spec.pcr=spec.pcr,
               sens.dpp=sens.dpp,
               spec.dpp=spec.dpp,
               prev=prev,
               prec.lpi=100
  )
  
  ## Initialize all prior values
  inits <- function(){
    list(lpi=rnorm(1,log(prev)/(1-log(prev)), 100),
         b1=rnorm(0,1),
         b2=rnorm(0,1),
         b3=rnorm(0,1),
         eps=rnorm(nind,0,5000))
  }
  
  ## Start MCMC simulation
  sens_spec.sim_m1 <- bugs(data=data, 
                           inits=inits, ## FIX THIS LINE
                           model.file="Model1_separate_tests_GB.txt",
                           parameters=c("b1", "b2", "b3","lpi", 
                                        "D", 
                                        "eps"),
                           n.chains=3, n.iter=30000, n.burnin=10000, 
                           codaPkg=TRUE, debug=TRUE)
  
  ## Use coda to read things back into R
  codaobject_m1 <- read.bugs(sens_spec.sim_m1)
  model1_df <- do.call(rbind.data.frame, codaobject_m1)
```
  
### OpenBUGS Model 1 Posterior Distributions
  
```{r}
par(mfrow=c(2,2))
## Graphial summaries of posterior distributions
hist(exp(model1_df$lpi)/(1+exp(model1_df$lpi)), main="prevalence", xlab="posterior draws")
abline(v=mean(exp(model1_df$lpi)/(1+exp(model1_df$lpi))), lty="dashed", col="red")
  
hist(model1_df$b1, main="Age", xlab="posterior draws")
abline(v=mean(model1_df$b1), lty="dashed", col="red")
hist(model1_df$b2, main="Sex", xlab="posterior draws")
abline(v=mean(model1_df$b2), lty="dashed", col="red")
hist(model1_df$b3, main="Age/Sex Interaction", xlab="posterior draws")
abline(v=mean(model1_df$b3), lty="dashed", col="red")
```
  
### OpenBUGS Model 1 Disease State Classification
  
```{r}
## Set up storage for model results
pred_df_m1 <- data.frame(obs=1:nind,
                         D=rep(NA,nind), ## average estimate
                         SD=rep(NA,nind),
                         LB=rep(NA,nind), ## 2.5th percentile
                         UB=rep(NA,nind), ## 97.5th percentile
                         model_assignment=rep(NA,nind),
                         Clinical_status=ss_data2$ClinicalStatus,
                         Diagnostic_status=ss_data2$Diagnostically_positive)
  
## Calculate probabilities of compartment membership for each posterior draw
pred_df_m1$D <- apply(model1_df[,grep("D", names(model1_df))], 2, mean)
pred_df_m1$SD <- apply(model1_df[,grep("D", names(model1_df))], 2, sd)
pred_df_m1$LB <- apply(model1_df[,grep("D", names(model1_df))], 2, 
                       quantile, probs=0.025)
pred_df_m1$UB <- apply(model1_df[,grep("D", names(model1_df))], 2, 
                         quantile, probs=0.975)
  
  
summary(pred_df_m1)
  
## Apply a cut off of point estimate of 0.5; if pi.D > 0.5, classify as S (symptomatic), otherwise as N;
## Summarize in a table (clinical status versus diagnostic status)
table(pred_df_m1[pred_df_m1$D > 0.5,]$Clinical_status, 
      pred_df_m1[pred_df_m1$D > 0.5,]$Diagnostic_status)
  
## Print summary table of clinical status versus diagnostic status from the original data
table(pred_df_m1$Clinical_status, pred_df_m1$Diagnostic_status)
  
  ## boxplots
p1 <- (ggplot(data=pred_df_m1, aes(x=Diagnostic_status, y=D)) 
       + geom_boxplot()
       + theme_bw())
p2 <- (ggplot(data=pred_df_m1, aes(x=Clinical_status, y=D)) 
       + geom_boxplot()
       + theme_bw())
  
ggarrange(p1,p2, nrow=1)
```
  
### OpenBUGS Model 1 Implementation - Informative Prior Prevalence
  
```{r include=FALSE, message=FALSE}
  ## Specify data
  y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
  y.dpp <- 1*(ss_data2$DPP_positive2=="Positive")
  
  nind <- nrow(ss_data2)
  
  Sex2 <- 1*(ss_data2$Sex=="M")
  Age <- as.numeric(ss_data2$Age)
  
  prev2 <- 0.3
  
  data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp=y.dpp,
               Sex=Sex2,
               Age=Age,
               sens.pcr=sens.pcr,
               spec.pcr=spec.pcr,
               sens.dpp=sens.dpp,
               spec.dpp=spec.dpp,
               prev=prev2,
               prec.lpi=100
  )
  
  ## Initialize all prior values
  inits <- function(){
    list(lpi=rnorm(1,log(prev2)/(1-log(prev2)), 10),
         b1=rnorm(0,1),
         b2=rnorm(0,1),
         b3=rnorm(0,1),
         eps=rnorm(nind,0,5000))
  }
  
  ## Start MCMC simulation
  sens_spec.sim_m1b <- bugs(data=data, 
                            inits=inits, ## FIX THIS LINE
                            model.file="Model1_separate_tests_GB.txt",
                            parameters=c("b1", "b2", "b3","lpi", 
                                         "D", 
                                         "eps"),
                            n.chains=3, n.iter=30000, n.burnin=10000, 
                            codaPkg=TRUE, debug=TRUE)
  
  ## Use coda to read things back into R
  codaobject_m1b <- read.bugs(sens_spec.sim_m1b)
  model1b_df <- do.call(rbind.data.frame, codaobject_m1b)
  ```
  
  ### OpenBUGS Model 1 Posterior Distributions - Informative Prior Prevalence
  
  ```{r}
  par(mfrow=c(2,2))
  ## Graphial summaries of posterior distributions
  hist(exp(model1b_df$lpi)/(1+exp(model1b_df$lpi)), main="prevalence", xlab="posterior draws")
  abline(v=mean(exp(model1b_df$lpi)/(1+exp(model1b_df$lpi))), lty="dashed", col="red")
  
  hist(model1b_df$b1, main="Age", xlab="posterior draws")
  abline(v=mean(model1b_df$b1), lty="dashed", col="red")
  hist(model1b_df$b2, main="Sex", xlab="posterior draws")
  abline(v=mean(model1b_df$b2), lty="dashed", col="red")
  hist(model1b_df$b3, main="Age/Sex Interaction", xlab="posterior draws")
  abline(v=mean(model1b_df$b3), lty="dashed", col="red")
```
